---
title: "Class 12"
author: "Gonzalez A16745338"
format: gfm

---

# Section 1. Proportion og G/G in a population


Downloaded a CSV file from Ensemble

Here we read this CSV file

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)


```


```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

Now lets look at a different population. I picked the GBR.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

head(gbr)
```

Find proportion of G G 

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) *100,2)
```

this varient that is associated with childhood asthma is more frequent in the GBR population the the MKL population.

Now letsdig further...



> Q13: Read this file into R and determine the sample size for each genotype and
their corresponding median expression levels for each of these genotypes. 
Hint: The read.table(),summary() and boxplot() functions will likely be useful
here. There is an example R script online to be used ONLY if you are struggling 
in vein. Note that you can find the medium value from saving the output of the 
boxplot() function to an R object and examining this object.There is also the 
medium() and summary() function that you can use to check your understanding.


```{r}
data <- read.table("rs8067378_ENSG00000172057.6.txt", header = TRUE)

summary(data)

library(dplyr)

summary_data <- data %>%
  group_by(geno) %>%
  summarise(
    Sample_Size = n(),
    Median_Expression = median(exp, na.rm = TRUE)
  )

print(summary_data)


```




> Q14: Generate a boxplot with a box per genotype, what could you infer from the
relative expression value between A/A and G/G displayed in this plot? Does the 
SNP effect the expression of ORMDL3? Hint: An example boxplot is provided 
overleaf – yours does not need to be as polished as this one. 


```{r}
data <- read.table("rs8067378_ENSG00000172057.6.txt", header = TRUE, sep = " ", na.strings = c("", "NA"))
str(data)
head(data)
summary(data)
```


```{r}
colnames(data) <- c("sample", "geno", "exp")
```


```{r}
# Filter out rows with NA 
data_clean <- data[!is.na(data$exp) & is.finite(data$exp), ]
```


```{r}
# boxplot 
boxplot(exp ~ geno, data = data_clean,
        main = "Expression Levels by Genotype",
        xlab = "Genotype", ylab = "Expression Levels",
        col = c("blue", "orange", "purple"))

```











