---
title: "Class 05: Data Visualization with GGPLOT"
author: "Gonzalez A16745338"
format: gfm
---

# Week 4 Data Visualization Lab

# Install the package ggplot2
# install.packages("ggplot2")

# Anytime I want to use this package I need to load it

```{r}
library(ggplot2)
#View(cars)
```


# A quick base R plot - this is not a ggplot2

```{r}
plot(cars)
```

# Our first ggplot
# We need data + aes + geoms

```{r}
ggplot(data=cars) +
  aes(x=speed, y=dist) +
  geom_point()
p <- ggplot(data=cars) +
      aes(x=speed, y=dist) +
      geom_point()
```
  


# Add a line geom with geom_line()
```{r}
p + geom_line()
```

# Add a trend line close to the data

```{r}
p + geom_smooth()
p + geom_smooth(method= "lm")
```



# Read in our drug expression data
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```



# Q. How many genes are in this dataset

```{r}
nrow(genes)
ncol(genes)
```



# Q. how many 'up' regulated genes

```{r}
table(genes$State)
```

# Q. What fraction of genes is up-regulated

```{r}
round((table(genes$State) / nrow(genes)) * 100, 2)
```


# Let's make a first plot attempt

```{r}
g <- ggplot(data=genes) +
      aes(x=Condition1, y=Condition2, col=State) +
      geom_point()
g
```
      


# Add some color

```{r}
g + scale_color_manual(values=c("blue", "gray", "red")) +
  labs(title = "Gene Expression changes", 
  x ="Control (no drug)") +
  theme_bw()
```












