---
title: "Class 16"
author: "Gonzalez A16745338"
format: pdf
---

This is a simple text editor called unix
Same key unix cmds so far

pwd: print working dir
cd: change
 dir
ls: list out files
mkdir: make a new dir/folder
nano: a basic text editor

  

cp notes.txt mycopy.txt
ls
mv mycopy.txt somesillyname.txt
ls
^q


ssh -i "~/Downloads/bimm143_gag002.pem" ubuntu@ec2-35-165-223-67.us-west-2.compute.amazonaws.com

scp -i "~/Downloads/bimm143_gag002.pem" ubuntu@ec2-35-165-223-67.us-west-2.compute.amazonaws.com


scp -i "~/Downloads/bimm143_gag002.pem" ubuntu@ec2-35-165-223-67.us-west-2.compute.amazonaws.com:~/work/mm-second.x.zebrafish.tsv myresult.tsv


```{r}
c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore")
```

```{r}
## Asuming your blast results are stored in an object called 'b'
# plot(b$pident  * (b$qend - b$qstart), b$bitscore)
```




```{r}
library(ggplot2)
# ggplot(b, aes(pident, bitscore)) + geom_point(alpha=0.1) 
```


```{r}
# ggplot(b, aes((b$pident * (b$qend - b$qstart)), bitscore)) + geom_point(alpha=0.1) + geom_smooth()
```






