---
title: "class 14"
author: "Gonzalez A16745338"
format: gfm
---

Today we will complete an RNASeq analysis from counts to pathways.

We will work with data on differential analysis of lung fibroblasts in response to loss of developmental transciption factor HOXA1.


## Data Import
```{r}
library(DESeq2)
```



```{r}
metaFile <- "GSE37704_metadata.csv"
countFile <- "GSE37704_featurecounts.csv"

# Import metadata and take a peak
metadata = read.csv(metaFile, row.names=1)
metadata$id <- row.names(metadata)
head(metadata)

```
```{r}
# Import countdata
countData = read.csv(countFile, row.names=1)
head(countData)
```

Check correspondence of the columns of 'counts' and rows of 'metadata'

```{r}
colnames(countData)
```

```{r}
metadata$id
```

We need to remove the first length column from our counts 

```{r}
counts <- countData[,-1]
colnames(counts) == metadata$id
```

> Q. How many genes do we have? 

```{r}
nrow(counts)
```

There are tones of zero count genes that we can get rid of.

```{r}
to.rm <- rowSums(counts) == 0
#counts[!to.rm,]
```
```{r}
to.keep <- rowSums(counts) == !0
countData[to.keep,]
```


> Q. How many genes do we have left?

```{r}
nrow(counts)
```


## DESeq setup
```{r}
library(DESeq2)
```


## DESeq Analysis


```{r}
dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = metadata, 
                              design = ~condition)
dds
```

```{r}
dds <- DESeq(dds)
res <- results(dds)
```



```{r}
head(res)
```


## Result Visualization

```{r}
plot( res$log2FoldChange, -log(res$padj) )

```












## Add Annotataion

```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)

columns(org.Hs.eg.db)
```


```{r}
res$symbol <- mapIds(org.Hs.eg.db,
              keys=rownames(res),
              keytype="ENSEMBL",
              column="SYMBOL")
```

```{r}
res$entrez = mapIds(org.Hs.eg.db,
                    keys=rownames(res),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

```



```{r}
head(res, 10)
```



## Pathway Analysis

```{r}
library(gage)
library(gageData)
library(pathview)
```

```{r}
data(kegg.sets.hs)
data(sigmet.idx.hs)

```


```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```


```{r}
# Get the results
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

```{r}
head(keggres$less, 3)
```


```{r}
head(keggres$greater, 3)
```


```{r}
nrow(countData)
nrow(res)
```



```{r}
sig_genes <- res[res$padj <= 0.05 & !is.na(res$padj), "symbol"]
print(paste("Total number of significant genes:", length(sig_genes)))
```

```{r}
write.table(sig_genes, file="significant_genes.txt", row.names=FALSE, col.names=FALSE, quote=FALSE)
```



## Save Results

```{r}
write.csv(res, file="myresults.csv")
```







